Mahotas: Open source software for scriptable computer vision

### 2012-08-22 paper.tex
MIN This bears saying
ENH More details

\title{Mahotas: Open source software for scriptable computer vision}
\author{Luis Pedro Coelho}

@@

Mahotas is a computer vision library for the Python. It operates on numpy
arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
by that project for storing information.

@@

Version 1.0 has been released recently and this is now a mature, well-tested
package (the first version was made available over 4~years ago). \footnote{Note
for reviewers: the version currently available implements all functionality
described in this manuscript. I will release it as version 1.0 when this
manuscript is accepted. Naturally, any bugs that are reported in the
meanwhile will be addressed.}

@@

\begin{description}
\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}.
\item[Wavelet] Haar and Daubechies wavelets.
\item[features] Some feature descriptors. In particular, Haralick texture
features, Zernike moments, local binary patterns, threshold adjacency
statistics.
\item[morphological functions] Erosion and dilation, as well as some more
complex operations.
\item[watershed] seeded watershed.
\item[polygon operations] convex hull, polygon drawing.
\end{description}

There are a few interface conventions which apply to many functions.

When meaningful, a structuring element is used to define neighbourhoods or
adjacency relationships (morphological functions, in particular, use this
convention). Generally, the default is to use a $3 \times 3$ cross as the
default if no structuring filter is given (the exception to this rule is the
median filter, where the default is a $3 \times 3$ square).

@@

Mahotas is written in \cpp, but almost always, the user calls a Python function
which checks types and then calls the internal function. This is slightly
slower, but it is easier to develop this way.

@@

>>>>

\title{Mahotas: Open source software for scriptable computer vision}
\author{Luis Pedro Coelho\\
Lane Center for Computational Biology, Carnegie Mellon University, Pittsburgh, USA\\
Instituto de Medicina Molecular, Lisboa, Portugal}

@@

Mahotas is a computer vision library for the Python. It operates on numpy
arrays~\citep{numpystructure}. Therefore, it uses all the infrastructure built
by that project for storing information. In particular, unlike libraries
written in the C~Language or in Java, Mahotas does not need to define a data
structure, but uses the numpy array. Many basic manipulation functionality that
would otherwise be part of a computer vision library are handled by numpy, for
example computing averages and other simple statistics, handling multi-channel
images, converting between types (integer and floating point images are
supported by mahotas, whenever it is meaningful). For the user, this has the
additional advantage that they do not need to learn yet another set of
functions.

On the other end, by integrating into the Python numeric ecosystem, users can
use other packages in a seamless way. In particular, mahotas does not implement
any machine learning functionality, but rather advises the user to use another,
specialised package, such as scikits-learn or milk.

@@

Version 1.0 of mahotas has been released recently and this is now a mature,
well-tested package (the first version was made available over 4~years ago).
\footnote{Note for reviewers: the version currently available implements all
functionality described in this manuscript. I will release it as version 1.0
when this manuscript is accepted to coincide with publication. Naturally, any
bugs that are found and reported in the meanwhile will be addressed.} Mahotas
runs and is used on different versions of Unix (including Linux, SunOS, and
FreeBSD), Mac OS X, and Windows.

@@

\begin{description}
\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This is
a collection of local features.
\item[features] Global feature descriptors. In particular, Haralick texture
features, Zernike moments, local binary patterns, threshold adjacency
statistics.
\item[Wavelet] Haar and Daubechies wavelets. Forward and reverse transforms.
\item[morphological functions] Erosion and dilation, as well as some more
complex operations built on these. There are both binary and grayscale
implementations of these operators.
\item[watershed] seeded watershed and distance map transforms.
\item[filtering] Gaussian filtering and general convolutions.
\item[polygon operations] convex hull, polygon drawing.
\end{description}

In general, functions operate on all data types (except when it not well
defined). This is performed without any extra memory copies. Mahotas is heavily
optimised for both speed and memory usage.

There are a few interface conventions which apply to many functions. When
meaningful, a structuring element is used to define neighbourhoods or adjacency
relationships (morphological functions, in particular, use this convention).
Generally, the default is to use a $3 \times 3$ cross as the
default if no structuring filter is given (the exception to this rule is the
median filter, where the default is a $3 \times 3$ square).

@@

Mahotas is written in \cpp, but this is completely hidden from the user as
there are hand-written Python wrappers for all functions (automatically
generated wrappers inevitably lead to worse error messages and are less
flexible).

@@

Mahotas uses the standard Python build machinery and distribution channels.
Building and installing from source code is done using
\begin{verbatim}
python setup.py install
\end{verbatim}
Alternatively, Python based package managers (such as \texttt{easy\_install} or
\texttt{pip}) can be used (mahotas works well with these systems).
